---
title: "simpleRegress"
author: "Rabi Elkhoury"
date: "Summer 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```

The plot that relates the variables LENGTH with NECK has a higher adjusted R-squared value (.7423) as compared to the relationship between the LENGTH and WEIGHT variables.  Additionally, the former yields a smaller residual standard error value (5.376), while the latter carried 5.432 degrees of freedom.  When considering this data, the variable NECK seems to be a more accurate predictor of LENGTH than WEIGHT.  


```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

The relative residual standard error and adjusted r-squared values are too high and low, respectively, as compared to the aforementioned variables, indicating the relationship between AGE and LENGTH is weak.   

```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

The scatter-plot detailing the relationship between LENGTH and HEADWTH, when computed, showed a higher residual standard error and lesser adjusted r-squared as compared to the scatter-plot between LENGTH and NECK.  


```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

The residual standard error (4.235) and adjusted r-squared value (.8434) relating LENGTH and HEADLEN better describes/predicts bear length compared to the scatter-plot illustrating the relationship between LENGTH and NECK, since the latter's residual standard error is lower, and adjusted r-squared value is higher.  

## Conclusion
In this introduction to scatter-plot and analysis, we were able to investigate the relationship among multivariate data, to determine the relative association between certain variables.  The primary method for comparison involved checking the residual standard error and adjusted r-value metrics.  Because the scatter-plot depicting the relation between the LENGTH and HEADLEN variables yielded the lower residual error and highest multiple r-squared values, it is reasonable to assume that the association between general bear length and bear head length is strongest.
